---
title: "Part 2: Interactive Graphics"
author: Max Kuhn (RStudio)
output:
  xaringan::moon_reader:
    css: ["mtheme_max.css", "fonts_mtheme_max.css"]  
    self_contained: false
    lib_dir: libs
    nature:
      ratio: '16:9'
      highlightLanguage: R
      countIncrementalSlides: false   
---


```{r startup, include=FALSE}
library(ggplot2)
knitr::opts_chunk$set(
  message = FALSE, 
  warning = FALSE
)
```



## These Slides

Can be found at [https://github.com/topepo/ncb-2019-shiny](https://github.com/topepo/ncb-2019-shiny).

---

# Towards interactivity

R has made big improvements on their graphics capabilities with [`lattice`](https://cran.r-project.org/package=lattice) and [`ggplot2`](https://cran.r-project.org/package=ggplot2)

However, these are _static_ graphics and could be improved with interactive capabilities. 

With **shiny**, it's actually possible to layer interaction on top of static R plots! Let's consider a small example.

---

# Example Data

We have data on some compounds measured under different conditions:

 * **type**: screening (via single point conc) vs confirmation (via IC50 dose-response)
 * **target** format: the drug target

```{r, message = FALSE}
library(ggplot2)
library(readr)
assays <- read_csv("IC50_data.csv")
str(assays)
```

We will plot the screening result versus the confirmation IC50 for both formats. If we only have the screening results, how have those related to the more definitive IC50?


---

# Screening results

.pull-left[
```r
p <- ggplot(
  data = assays, 
    aes(
      x = Single_Point, 
      y = Dose_Response, 
      color = Target
    )
  ) +
  geom_point(size = 4, alpha = .4) +
  geom_smooth(se = FALSE) +  
  scale_y_log10()
p
```

What about that point with a low `Dose_Response`? What compound is it?
]


.pull-right[
```{r, echo = FALSE}
p <- ggplot(data = assays, aes(x = Single_Point, y = Dose_Response, color = Target)) +
  geom_point(size = 4, alpha = .4) +
  geom_smooth(se = FALSE) +  
  scale_y_log10()
p
```
]



---

## Graphical queries in shiny

```r
library(shiny)

ui <- fluidPage(
  # Exposes hover info on the server as `input$hover`
  # Can also do this with click or brush
  plotOutput("plot", hoverId = "hover"),
  tableOutput("table")
)

server <- function(input, output, session) {
  output$plot <- renderPlot(p)
  output$table <- renderTable(nearPoints(assays, input$hover))
}

shinyApp(ui, server)
```

Learn more about plot interactions in **shiny** <https://shiny.rstudio.com/articles/#graphics>

---

## Getting more interactive

**shiny**'s static plot interaction is great for getting user input from one view and sending it to another (e.g. displaying a table of selected point(s)).

Sometimes it pays off if the graphic itself is web-based and interactive. Many R packages can produce these via the **htmlwidgets** package, which provides a consistent interface across different contexts (e.g., R console, RStudio, shiny, rmarkdown, etc)
  * To see a gallery of the most popular htmlwidgets (we'll cover some of these) -- https://gallery.htmlwidgets.org/
  * To learn more about htmlwidgets itself -- http://www.htmlwidgets.org/
    
Today we'll introduce two interactive graphing libraries: **ggiraph** and **plotly**. 
  * If you know **ggplot2**, you can easily use these packages to make your static graphs interactive.
  * Due to their different approaches, **ggiraph** does a better job of maintaining the look and feel of the **ggplot2** graph, but **plotly** provides a more fully-featured and extensible framework (you can also use it without **ggplot2**).
  
---
# `plotly`

It's easiest to get started with **plotly**'s `ggplotly()`. This function converts a static **ggplot2** to an interactive web-based graphic.

.pull-left[

```r
library(plotly)
p <- ggplot(data = assays, aes(x = Single_Point, y = Dose_Response, color = Target)) +
  geom_point(size = 4, alpha = .4) +
  geom_smooth(se = FALSE) +  
  scale_y_log10()
ggplotly(p)
```

]

.pull-right[
<iframe src="ggplotly.html" width="100%" height=500 scrolling="no" seamless="seamless" frameBorder="0">
]

  
---
# `plotly`

It's easiest to get started with **plotly**'s `ggplotly()`. This function converts a static **ggplot2** to an interactive web-based graphic.

.pull-left[

```r
library(plotly)
p <- ggplot(data = assays, aes(x = Single_Point, y = Dose_Response, color = Target)) +
  geom_point(aes(text = Compound), size = 4, alpha = .4) +
  geom_smooth(se = FALSE) +  
  scale_y_log10()
# show only the text aes (instead of all of them) in the tooltip
ggplotly(p, tooltip = "text")
```

To control the tooltip that appears, use the `text` aesthetic (notice how `Compound` is now shown on hover).

]

.pull-right[
<iframe src="ggplotly-text.html" width="100%" height=500 scrolling="no" seamless="seamless" frameBorder="0">
]


---
# `plotly`

You can also make **plotly** graphs without **ggplot2**, which is useful for stuff like [3D charts](https://plotly-r.com/d-charts.html), [maps](https://plotly-r.com/maps.html), and generally having more control. Read more at <https://plotly-r.com> 

.pull-left[

```{r, eval = FALSE}
plot_ly(
  assays, x = ~Single_Point, y = ~Dose_Response, 
  color = ~Target, text = ~Compound,
  size = I(50), alpha = 0.4
) %>% layout(yaxis = list(type = "log"))
```

**plotly** also provides the building block for other useful packages like [**heatmaply**](https://cran.r-project.org/web/packages/heatmaply/vignettes/heatmaply.html) and [**TSStudio**](https://cran.r-project.org/web/packages/TSstudio/index.html)
]

.pull-right[
<iframe src="plotly.html" width="100%" height=500 scrolling="no" seamless="seamless" frameBorder="0">
]


---

# Access plotly events in shiny

```r
library(shiny)

ui <- fluidPage(
  plotlyOutput("plot"),
  tableOutput("table")
)

server <- function(input, output, session) {
  output$plot <- renderPlotly({
    p <- ggplot(assays, aes(x = Single_Point, y = Dose_Response, color = Target)) +
      geom_point(aes(text = Compound, customdata = Compound), size = 4, alpha = .4) +
      geom_smooth(se = FALSE) +  
      scale_y_log10()
    ggplotly(p, tooltip = "text")
  })
  # event_data() function can be used to access information about all sorts of events
  output$table <- renderTable({
    hovered_compound <- event_data("plotly_hover")$customdata
    filter(assays, Compound %in% hovered_compound)
  })
  
}

shinyApp(ui, server)
```

---

# Exercise

Run the following example apps

```r
plotly_example("shiny", "event_data")
plotly_example("shiny", "event_data_3D")
```

How many different plotly events are these apps showing? Can you trigger all of them?


---

# `ggiraph`

**plotly**'s `ggplotly()` doesn't always get the conversion right. **ggiraph** does a better at preserving the look, but doesn't have quite as many interactive features as **plotly**.

.pull-left[
```{r ggiraph_example, eval=FALSE}
library(ggiraph)
## We could add a column for click behavior:
# assays$click <- {a vector of urls}
ggiraph_ggplot <- 
  ggplot(assays, aes(x = Single_Point, y = Dose_Response, color = Target)) +
  geom_point_interactive(
    # could also add `onclick = click`
    aes(tooltip = Compound), 
    size = 4, 
    alpha = .4
  ) +
  geom_smooth(se = FALSE) +  
  scale_y_log10()

## show using
ggiraph(ggobj = ggiraph_ggplot, width = 1)
```

Note the use of `geom_point_interactive()` instead of `geom_point()`.
]


.pull-right[
```{r ggiraph_show, echo = FALSE}
library(ggiraph)
## We could add a column for click behavior:
# assays$click <- {a vector of urls}

ggiraph_ggplot <- 
  ggplot(assays, aes(x = Single_Point, y = Dose_Response, color = Target)) +
  geom_point_interactive(
    # could also add `onclick = click`
    aes(tooltip = Compound), 
    size = 4, 
    alpha = .4
  ) +
  geom_smooth(se = FALSE) +  
  scale_y_log10()

## show using
ggiraph(ggobj = ggiraph_ggplot, width = 1)
```
]

---

# Access ggiraph events in shiny

```r
library(shiny)
library(ggiraph)

ggiraph_ggplot <- 
  ggplot(assays, aes(x = Single_Point, y = Dose_Response, color = Target)) +
  geom_point_interactive(
    aes(tooltip = Compound, data_id = Compound), 
    size = 4, 
    alpha = .4
  ) +
  geom_smooth(se = FALSE) +  
  scale_y_log10()

ui <- fluidPage(
  girafeOutput("plot"),
  tableOutput("table")
)

server <- function(input, output, session) {
  output$plot <- renderGirafe({
    ggiraph(
      ggobj = ggiraph_ggplot, 
      hover_css = "fill:red;cursor:pointer;",
      selection_type = "single",
      selected_css = "fill:red;"
    )
  })
  
  output$table <- renderTable({
    filter(assays, Compound %in% input$plot_selected)
  })
}

shinyApp(ui, server)
```


---

# Where to Find More

* More about **shiny**'s "static" interactivity <https://shiny.rstudio.com/articles/#graphics>
* More about **plotly** <https://plotly-r.com/>
* More about **ggiraph** <http://davidgohel.github.io/ggiraph/>
* More about **htmlwidgets** <http://www.htmlwidgets.org/>