---
title: "Part 2: Interactive Graphics"
author: Max Kuhn (RStudio)
output:
  xaringan::moon_reader:
    css: ["mtheme_max.css", "fonts_mtheme_max.css"]  
    self_contained: false
    lib_dir: libs
    nature:
      ratio: '16:9'
      highlightLanguage: R
      countIncrementalSlides: false   
---


```{r startup, include=FALSE}
library(ggplot2)
```



## These Slides

Can be found at [https://github.com/topepo/ncb-2019-shiny](https://github.com/topepo/ncb-2019-shiny).

---

# Better Graphics

R has made big improvements on their graphics capabilities with [`lattice`](https://cran.r-project.org/package=lattice) and [`ggplot2`](https://cran.r-project.org/package=ggplot2)

However, these are _static_ graphics and could be improved with interactive capabilities. 

At the same time, there are a large number of Javascript libraries that enable interactions:

 * [Data Driven Documents](http://d3js.org) (aka `D3`)
 * [BonsaiJS](https://bonsaijs.org/)
 * [`Chart.js`](http://www.chartjs.org)
 * [`n3`](http://n3-charts.github.io/line-chart/#/home)
 * [highcharts](http://www.highcharts.com)
 * [plot.ly](https://plot.ly/javascript/)


---

# R Interfaces

How can we blend these with R (and `shiny`)?

There are a good number of R packages that can do this:

 * [`ggvis`](https://cran.r-project.org/package=ggvis) and [`ggiraph`](https://cran.r-project.org/package=ggiraph) are designed specifically to compliment [`ggplot2`](https://cran.r-project.org/package=ggplot2)
 * [`rCharts`](https://github.com/ramnathv/rCharts) is a package that can create graphics using a few different JS libraries (GitHub only; maybe deprecated)
 * [`rVega`](https://github.com/metagraf/rVega) is a wrapper to the [Vega](http://vega.github.io/vega/) JS library, [`vegalite`](https://github.com/hrbrmstr/vegalite) too. 
 * [`networkD3`](https://cran.r-project.org/package=networkD3) can show interactive network diagrams 
 * [`plotly`](https://cran.r-project.org/package=plotly) works with [plot.ly](https://plot.ly/javascript/)
 * [`rbokeh`](https://github.com/bokeh/rbokeh) works with the [Bokeh Python interactive visualization library](http://bokeh.pydata.org/en/latest/) 
 * Many others, including [`metricsgraphics`](https://cran.r-project.org/package=metricsgraphics), [`scatterD3`](https://cran.r-project.org/package=scatterD3), [`threejs`](https://cran.r-project.org/package=threejs), [`heatmaply`](https://cran.r-project.org/web/packages/heatmaply/index.html), etc.
 
 
 Examples follow with application directories
 
---

# Example Data

We have data on some compounds measured under different conditions:

 * **type**: screening (via single point conc) vs confirmation (via IC50 dose-response)
 * **target** format: the drug target

```{r}
library(ggplot2)
library(readr)
assays <- read_csv("IC50_data.csv")
str(assays)
```

We will plot the screening result versus the confirmation IC50 for both formats. If we only have the screening results, how have those related to the more definitive IC50?


---

# Let's Make This Plot (or One Like It)

```{r, example, message = FALSE}
thm <- theme_bw() + 
  theme(
    panel.background = element_rect(fill = "transparent", colour = NA), 
    plot.background = element_rect(fill = "transparent", colour = NA),
    legend.position = "top",
    legend.background = element_rect(fill = "transparent", colour = NA),
    legend.key = element_rect(fill = "transparent", colour = NA)
  )
theme_set(thm)

base_plot <- 
  ggplot(data = assays, aes(x = Single_Point, y = Dose_Response, color = Target)) +
  geom_smooth(se = FALSE) +  
  scale_y_log10()

# We'll use different variants of `geom_point` in other slides
basic_ggplot <- base_plot + geom_point( size = 4, alpha = .4) 
```


---

# `ggiraph`

I like this one best of all since the syntax is almost identical to `ggplot2`'s and it is very low maintenance


.pull-left[

```{r ggiraph_example, eval=FALSE}
## We could add a column for click behavior:
# assays$click <- {a vector of urls}

ggiraph_ggplot <- 
  base_plot + 
  geom_point_interactive(
    # could also add `onclick = click`
    aes(tooltip = Compound), 
    size = 4, 
    alpha = .4
  ) 

## show using
ggiraph(ggobj = ggiraph_ggplot, width = 1)
```

]
.pull-right[

```{r ggiraph_show,echo = FALSE, eval = TRUE, message = FALSE}
library(ggiraph)
## We could add a column for click behavior:
# assays$click <- {a vector of urls}

ggiraph_ggplot <- 
  base_plot + 
  geom_point_interactive(
    # could also add `onclick = click`
    aes(tooltip = Compound), 
    size = 4, 
    alpha = .4
  ) 

## show using
ggiraph(ggobj = ggiraph_ggplot, width = 1)
```

]

---

# `ggiraph` in `shiny`: 

```{r ggiraph_shiny, eval=FALSE}
## in server.R
output$that_plot <- renderggiraph(
  ## code to make object `p` here
  ggiraph(ggobj = p, width = 1, width_svg = 5, height_svg = 3)
  )

## in ui.R
ggiraphOutput("that_plot")
```


---

# `plotly`

`plotly` started off as a web application that was inspired by `ggplot` but released an R package as a stand-alone piece of software (i.e. it does not use their servers). 


.pull-left[

.code70[

```{r plotly-example, eval=FALSE}
library(plotly)
plot_ly(assays, x = Single_Point, y = Dose_Response, 
        text = paste("compound: ", Compound),
        mode = "markers", color = Target)

## or with ggplot syntax
p <- 
  base_plot +
  geom_point(aes(text = paste("Compound:", Compound)), 
             size = 4, alpha = .4) 

## convert to plot.ly format
ggplotly(p)
```

]

]
.pull-right[
```{r plotly-show, results='asis', comment = NA, message=FALSE,warning=FALSE, echo = FALSE, eval = FALSE}
unloadNamespace(ggiraph)
library(plotly)
p <- 
  base_plot + 
  geom_point(aes(text = paste("Compound:", Compound)), 
             size = 4, alpha = .4) 
ggplotly(p)
```
]

---

# `plotly` in `shiny` 

```{r plotly_shiny, eval=FALSE}
## in server.R
p <- 
  base_plot + 
  geom_point(aes(text = paste("Compound:", Compound)), 
             size = 4, alpha = .4) 
output$that_plot <- renderPlotly(ggplotly(p))

## in ui.R
plotlyOutput("that_plot")
```

See the `plotly` example application. 

---

# JMP-Like Behavior with `plotly` 

The _latest_ version of `plotly` has hooks that allow the user to know which samples are selected at any given time. 

This information can be used to show the data, create new plots, highlight in other plots, etc. 

If you have at least `plotly_3.4.1`, you can see then example by running:

```{r plotly_events, eval=FALSE}
shiny::runApp(system.file("examples", "plotlyEvents", package = "plotly"))
```

[This page](https://plot.ly/r/shiny-coupled-events/) also shows examples. 


---

# `rbokeh`

This is a port of the Bokeh Python interactive visualization library to R. Once a figure is created, you can add points and lines to it. 

```{r rbokeh_example, eval=FALSE}
library(rbokeh) 
p <- figure()
p <- ly_points(fig = p, 
               x = Single_Point, y = Dose_Response, 
               data = assays,
               color = Target, hover = Compound) 
a_smooth <- with(subset(assays, Target == "A"),
                   lowess(Dose_Response ~ Single_Point))
b_smooth <- with(subset(assays, Target == "B"),
                   lowess(Dose_Response ~ Single_Point))   
p <- ly_lines(fig = p, a_smooth, legend = "A", color = "blue") 
p <- ly_lines(fig = p, b_smooth, legend = "B", color = "orange") 
y_axis(fig = p, log = TRUE)
```

---

# With Pipes!

.pull-left[

.code70[

```{r rbokeh_pipes, eval=FALSE}
library(rbokeh)
a_smooth <- with(subset(assays, Target == "A"),
                   lowess(Dose_Response ~ Single_Point))
b_smooth <- with(subset(assays, Target == "B"),
                   lowess(Dose_Response ~ Single_Point))  
p <- figure() %>% 
  ly_points(x = Single_Point,  y = Dose_Response, 
            data = assays, 
            color = Target, hover = Compound) %>%
  y_axis(log = TRUE) %>%
  ly_lines(a_smooth, legend = "A", color = "blue")  %>%
  ly_lines(b_smooth, legend = "B", color = "orange") 
p
```

]

]
.pull-right[


```{r rbokeh_show, results='asis', comment = NA, message=FALSE,warning=FALSE, echo = FALSE}
library(rbokeh)
a_smooth <- with(subset(assays, Target == "A"),
                   lowess(Dose_Response ~ Single_Point))
b_smooth <- with(subset(assays, Target == "B"),
                   lowess(Dose_Response ~ Single_Point))  
p <- figure() %>% 
  ly_points(x = Single_Point,  y = Dose_Response, 
            data = assays, 
            color = Target, hover = Compound) %>%
  y_axis(log = TRUE) %>%
  ly_lines(a_smooth, legend = "A", color = "blue")  %>%
  ly_lines(b_smooth, legend = "B", color = "orange") 
p
```

]





---

# `rbokeh` in `shiny`: 

```{r rbokeh_shiny, eval=FALSE}
## in server.R
output$new_scatter <- renderRbokeh({
  p <- figure()
  p <- ly_points(fig = p, 
                 x = Single_Point, y = Dose_Response, 
                 data = assays,
                 color = Target, hover = Compound) 
  ## and so on
  p
}) ## output$new_scatter  

## in ui.R
rbokehOutput("new_scatter")
```

---

# Where to Find More

The CRAN Task view on [Web Technologies](https://cran.r-project.org/web/views/WebTechnologies.html) has a curated list of `shiny` and javascript visualization packages (and more)
  
.font60[

```{r}
sessionInfo()
```

]